### 2025-08-16 basic-statistics.qmd
20250816 - comparing mixed models to ANOVA/multiple regression

Thus, in most cases, the regression framework can be used to examine most questions regarding associations between variables or differences between groups.

@@

>>>>

Thus, in most cases, the regression framework can be used to examine most questions regarding associations between variables or differences between groups.
Indeed, [*t*-test](#sec-tTest), [ANOVA](#sec-anova), and [regression](#sec-regression) are all special cases of the general linear model, which in turn, is a special case of [mixed models](#sec-mixedModels) without [random effects](#sec-mixedModelsOverview) [@Brauer2018].

@@

ANOVA has key limitations.
ANOVA cannot handle continuous [independent](#sec-experiment) or [predictor](#sec-correlationalStudy) variables that vary within units [@Brauer2018].
In addition, ANOVA poorly handles missing data—it uses listwise deletion and thus excludes participants with missing values on any of the [independent](#sec-experiment) or [dependent](#sec-experiment) variables [@Brauer2018].

### 2025-08-16 mixed-models.qmd
20250816 - comparing mixed models to ANOVA/multiple regression

When including random effects of a variable (e.g., age) in a mixed model, it is also important to include fixed effects of that variable in the model.
This is because random effects have a mean of zero.
Fixed effects allow the mean to differ from zero.
Thus, inclusion of random effects without the corresponding fixed effect can lead to bias in estimation of the association between the predictor variables and the outcome variable.

@@

Examples of Simpson's Paradox are depicted in Figures [-@fig-simpsonParadox] and [-@fig-simpsonsParadox].
In the example below, there is a positive association between the predictor variable and outcome variable for each group.
However, when the groups are combined, there is a *negative* association between predictor variable and outcome variable.
That is, the sign of an association can differ at different levels of analysis (e.g., group level versus person level).

@@

Consider that we observe a between-person association between a predictor, for example, how much sports drink a player drinks and their performance, such that players who drink more sports drink before games tend to perform better during games.
However, based on this association, if we draw the inference that sports drink consumption leads to better performance, this could be a faulty inference.
It is possible that, at the *within*-person level, there is no association or even a negative association between sports drink consumption and performance.
For helping to approximate causality, a much stronger test than relying on the between-person association would be to examine the association within the individual.
That is, examining the association within the individual would examine: when a player consumes more sports drink, whether they perform better than when the same player consumes less sports drink.
We describe within-person analyses to approximate causal inferences in @sec-causalInferenceWithinSubject.

In short, we often need to account for the groups or "levels" within which people are nested.
When multiple observations are nested within the same upper-level unit, they are often correlated.
That is, some variance is attributable to the lower-level unit (e.g., the player) and some variance is attributable to the upper-level unit (e.g., the team).
It is important to evaluate how much variance is attributable to the upper-level unit.
A way of evaluating how much variance is attributed to an upper-level unit is with the intraclass correlation coefficient (ICC).
If substantial variance is attributable to the upper-level unit, it is important to account for the nested structure of the data.
Mixed models are a useful approach to account for data with a nested structure so you can avoid committing the ecological fallacy.

@@

For fitting mixed models, the data should be in [long form](#sec-longWideForm)—each player (or unit of interest) should have multiple rows based on the grouping structure.
When estimating mixed models, it is important to know the grouping structure of the data because the data structure influences the model syntax.
Examples of data from nonindependent units include nested data, multiple membership data, and cross-classified data, as depicted in @fig-nestedDataStructures.

@@

Another form of nonindependent data is when data are *cross-classified* (also called crossed).
Data are cross-classified when the lower-level units are classified by two or more upper-level units, but the upper-level units are not hierarchical or nested within one another.
For instance, children may be nested within the crossing of schools and neighborhoods.
That is, children are nested within schools, and children are also nested within neighborhoods; however, children attending the same school may not necessarily be from the same neighborhood.
The data would still have a two-level hierarchy, but the data would be nested in specific school-neighborhood combinations.
That is, children are nested within the cross-classification of schools and neighborhoods.

Applied to longitudinal fantasy football data, Most simply, players' longitudinal performance data would be modeled as nested data with a two-level model: timepoints nested within players.
However, you could get more complicated.
For instance, you could estimate a three-level model of timepoints nested within players nested within teams.
This would help account for players' productivity benefitting or taking a hit from being on a given team, and it would account for players changing teams.
Alternatively, consistent with @Usami2018, you could estimate a cross-classified model of players' individual observations/measurement occasions belonging to two upper-level units: players and time points (e.g., week 1, week 2, 2025 season, etc.).
This would help account for systematic effects that are specific to a given timepoint (e.g., week or season).

@@

When trying to determine what predictors to include and how complex of a model to fit, it can be helpful to compare a variety a models to determine how well they fit the data (i.e., how much variance they account for).
To compare non-nested models, you can use fit criteria such as Akaike information criterion (AIC), which penalizes model complexity.
If two models are nested, you can compare their model fit using a likelihood ratio test.
Two models are considered nested if one model includes all of the terms of the original model along with additional terms.
The simpler model is considered "nested within" the more complex model.
In the likelihood ratio test, a significant difference indicates that the more complex model is significantly better fitting than the simpler model.
If the likelihood ratio test is non-significant, it indicates that the more complex model is not significantly better fitting, and thus you should prefer the more parsimonious model.

@@

In this chapter, we use mixed models to demonstrate how to model trajectories (growth curves) of players' fantasy points as a function of age.
One of the challenges when modeling players' fantasy points as a function of age is that fantasy points are a function of both ability and opportunity.
With age, the player's ability and opportunity may both decline, but it is difficult to disentangle the extent to which a player's decline is related to declines in ability versus opportunity.
As players get older, they may be supplanted by younger, more talented players.
With age, despite still having latent (unobserved) ability, players will get fewer and, eventually, no opportunity.
And we do not know the counterfactual—how many fantasy points they would have scored had they been given the full opportunity each season.
Thus, players may go from scoring 100+ points in a season to all of a sudden scoring way fewer points, with little in the way of intermediate steps.
Thus, this should inform how you interpret the resulting trajectories.
For instance, the steep declines can make it look like the model-implied fantasy points go well below zero at later ages, even though it is obviously not likely for players to obtain such a magnitude of negative fantasy points.
The negative model-implied values of fantasy points at later ages are likely an artifact of the decreasing opportunity that players tend to get as they age.
The crossover point, where the positive values becomes negative might indicate, for instance, that players at that age tend to have retired and thus have no opportunity.
Regardless, the form/shape/steepness of the decline is still potentially meaningful even if the precise negative number at a given age is not.

@@

Based on the [scatterplot](#sec-scatterplot) (and the bivariate association below), Quarterbacks' fantasy points appear to increase with age.

@@

Mixed models allow accounting for multiple levels or units of analysis and to include both fixed and random effects.
Inclusion of random effects allows the association between the predictor variables (the intercept and age) and the outcome variable (fantasy points) to differ for each individual in the grouping level (in this case, each player).
This allows for more accurately predicting phenomena.
Based on the bivariate scatterplots between age and fantasy points, we might conclude that players tend to stay stable or even increase in fantasy points with age.
However, this conclusion would be wrong.
When we account for the longitudinal data using mixed models, we observe that players' fantasy points tend to decrease with age, with the timing and rate of decline differing for each position.
In other words, the association between age and fantasy points differs at the person level versus the group level, which is an example of [Simpson's paradox](#sec-simpsonsParadox).
In sum, mixed models are valuable for examining associations between variables when there are multiple levels of data (i.e., multiple observations within the same unit, known as clustering or nesting).
It is important not to confuse the association at one level (e.g., group level) with the association at another level (e.g., person level), which is an example of the [ecological fallacy](#sec-fallaciesEcological).

>>>>

When including random effects of a variable (e.g., age) in a mixed model, it is also important to include fixed effects of that variable in the model.
This is because random effects have a mean of zero.
Fixed effects allow the mean to differ from zero.
Thus, inclusion of random effects without the corresponding fixed effect can lead to bias in estimation of the association between the [predictor variables](#sec-correlationalStudy) and the [outcome variable](#sec-correlationalStudy).

Mixed modeling is a more flexible modeling framework than [analysis of variance](#sec-anova) (ANOVA).
Unlike [ANOVA](#sec-anova), mixed modeling can accommodate both continuous and categorical [predictor variables](#sec-correlationalStudy) that vary within units [@Brauer2018].
[ANOVA](#sec-anova), cannot handle continuous [predictor variables](#sec-correlationalStudy) that vary within units [@Brauer2018].
In addition, [ANOVA](#sec-anova) yields biased inferences when the same participants are exposed to the same items or stimuli [@Brauer2018].
Additionally, mixed models better handle missing data compared to [ANOVA](#sec-anova) [@Brauer2018].

@@

Simpson's paradox occurs when the association between between the [predictor variable](#sec-correlationalStudy) and [outcome variable](#sec-correlationalStudy) for the subgroups differ from the association when the subgroups are combined.
Examples of Simpson's paradox are depicted in Figures [-@fig-simpsonParadox] and [-@fig-simpsonsParadox].
In the example below, there is a positive association between the [predictor variable](#sec-correlationalStudy) and [outcome variable](#sec-correlationalStudy) for each group.
However, when the groups are combined, there is a *negative* association between [predictor variable](#sec-correlationalStudy) and [outcome variable](#sec-correlationalStudy).
That is, the sign of an association can differ at different levels of analysis (e.g., group level versus person level).

@@

Consider that we observe a between-person association between a [predictor](#sec-correlationalStudy), for example, how much sports drink a player drinks and their performance, such that players who drink more sports drink before games tend to perform better during games.
However, based on this association, if we draw the inference that sports drink consumption leads to better performance, this could be a faulty inference.
It is possible that, at the *within*-person level, there is no association or even a negative association between sports drink consumption and performance.
For helping to approximate causality, a much stronger test than relying on the between-person association would be to examine the association within the individual.
That is, examining the association within the individual would examine: when a player consumes more sports drink, whether they perform better than when the same player consumes less sports drink.
We describe within-person analyses to approximate causal inferences in @sec-causalInferenceWithinSubject.

In short, we often need to account for the groups or "levels" within which people are nested.
When multiple observations are nested within the same upper-level unit, they are often correlated.
That is, some variance is attributable to the lower-level unit (e.g., the player) and some variance is attributable to the upper-level unit (e.g., the team).
It is important to evaluate how much variance is attributable to the upper-level unit.
A way of evaluating how much variance is attributable to an upper-level unit is with the intraclass correlation coefficient (ICC).
If substantial variance is attributable to the upper-level unit, it is important to account for the nested structure of the data.
Mixed models are a useful approach to account for data with a nested structure so you can avoid committing the ecological fallacy.

@@

For fitting mixed models, the data should be in [long form](#sec-longWideForm)—each player (or unit of interest) should have multiple rows based on the grouping structure.
The cluster variable is the variable or variables that represent the upper-level unit(s), within which the data are nested.
When estimating mixed models, it is important to know the grouping structure of the data because the data structure influences the model syntax.
Examples of data from nonindependent units include nested data, multiple membership data, and cross-classified data, as depicted in @fig-nestedDataStructures.

@@

Another form of nonindependent data is when data are *cross-classified* (also called crossed).
Data are cross-classified when the lower-level units are classified by two or more upper-level units, but the upper-level units are not hierarchical or nested within one another.
For instance, children may be nested within the crossing of schools and neighborhoods.
That is, children are nested within schools, and children are also nested within neighborhoods; however, children attending the same school may not necessarily be from the same neighborhood, and children from the same neighborhood may not necessarily attend the same school.
The data would still have a two-level hierarchy, but the data would be nested in specific school-neighborhood combinations.
That is, children are nested within the cross-classification of schools and neighborhoods.

Applied to longitudinal fantasy football data, Most simply, players' longitudinal performance data would be modeled as nested data with a two-level model: timepoints nested within players, with the player's ID as the cluster variable.
However, you could get more complicated.
For instance, you could estimate a three-level model of timepoints nested within players nested within teams.
This would help account for players' productivity benefitting or taking a hit from being on a given team, and it would account for players changing teams.
Alternatively, consistent with @Usami2018, you could estimate a cross-classified model of players' individual observations/measurement occasions belonging to two upper-level units: players and time points (e.g., week 1, week 2, 2025 season, etc.).
This would help account for systematic effects that are specific to a given timepoint (e.g., week or season).

## Handling of Missing Data {#sec-mixedModelMissingness}

Mixed models handle missing data better than [ANOVA](#sec-anova) and [multiple regression](#sec-multipleRegression) [@Brauer2018].
[ANOVA](#sec-anova) and [multiple regression](#sec-multipleRegression) use listwise deletion by default and, thus, exclude participants who have missing values on any of the variables in the model (i.e., [independent/dependent](#sec-experiment) or [predictor/outcome](#sec-correlationalStudy)).
By contrast, with the data in [long form](#sec-longWideForm), mixed models make use of all rows that have values on all of the [predictor variables](#sec-correlationalStudy).
That is, mixed models exclude only those rows that have a missing value on any of the [predictor variables](#sec-correlationalStudy).
Unlike [ANOVA](#sec-anova) and [multiple regression](#sec-multipleRegression), mixed models retain rows with missing values on the [outcome variables](#sec-correlationalStudy).

@@

When trying to determine what [predictors](#sec-correlationalStudy) to include and how complex of a model to fit, it can be helpful to compare a variety a models to determine how well they fit the data (i.e., how much variance they account for).
To compare non-nested models, you can use fit criteria such as Akaike information criterion (AIC), which penalizes model complexity.
If two models are nested, you can compare their model fit using a likelihood ratio test.
Two models are considered nested if one model includes all of the terms of the original model along with additional terms.
The simpler model is considered "nested within" the more complex model.
In the likelihood ratio test, a significant difference indicates that the more complex model is significantly better fitting than the simpler model.
If the likelihood ratio test is non-significant, it indicates that the more complex model is not significantly better fitting, and thus you should prefer the more parsimonious model.

@@

In this chapter, we use mixed models to demonstrate how to model trajectories (growth curves) of players' fantasy points as a function of age.
When estimating growth curve models, time points are nested within individuals (i.e., players).
Mixed models provide a flexible framework for estimating growth curve models because they allow unequal time intervals between measurement occasions, time intervals that differ between individuals, and different numbers of observations for each individual [@Brauer2018].

One of the challenges when modeling players' fantasy points as a function of age is that fantasy points are a function of both ability and opportunity.
With age, the player's ability and opportunity may both decline, but it is difficult to disentangle the extent to which a player's decline is related to declines in ability versus opportunity.
As players get older, they may be supplanted by younger, more talented players.
With age, despite still having latent (unobserved) ability, players will get fewer and, eventually, no opportunity.
And we do not know the counterfactual—how many fantasy points they would have scored had they been given the full opportunity each season.
Thus, players may go from scoring 100+ points in a season to all of a sudden scoring way fewer points, with little in the way of intermediate steps.
Thus, this should inform how you interpret the resulting trajectories.
For instance, the steep declines can make it look like the model-implied fantasy points go well below zero at later ages, even though it is obviously not likely for players to obtain such a magnitude of negative fantasy points.
The negative model-implied values of fantasy points at later ages are likely an artifact of the decreasing opportunity that players tend to get as they age.
The crossover point, where the positive values becomes negative might indicate, for instance, that players at that age tend to have retired and thus have no opportunity.
Regardless, the form/shape/steepness of the decline is still potentially meaningful even if the precise negative number at a given age is not.

@@

Based on the [scatterplot](#sec-scatterplot) (and the bivariate association below), Quarterbacks' fantasy points appear to (slightly) increase with age.

@@

Mixed models allow accounting for multiple levels or units of analysis and to include both fixed and random effects.
Inclusion of random effects allows the association between the [predictor variables](#sec-correlationalStudy) (the intercept and age) and the [outcome variable](#sec-correlationalStudy) (fantasy points) to differ for each individual in the grouping level (in this case, each player).
This allows for more accurately predicting phenomena.
Based on the bivariate scatterplots between age and fantasy points, we might conclude that players tend to stay stable or even increase in fantasy points with age.
However, this conclusion would be wrong.
When we account for the longitudinal data using mixed models, we observe that players' fantasy points tend to decrease with age, with the timing and rate of decline differing for each position.
In other words, the association between age and fantasy points differs at the person level versus the group level, which is an example of [Simpson's paradox](#sec-simpsonsParadox).
In sum, mixed models are valuable for examining associations between variables when there are multiple levels of data (i.e., multiple observations within the same unit, known as clustering or nesting).
It is important not to confuse the association at one level (e.g., group level) with the association at another level (e.g., person level), which is an example of the [ecological fallacy](#sec-fallaciesEcological).

### 2025-08-16 multiple-regression.qmd
20250816 - comparing mixed models to ANOVA/multiple regression

@@

where $p$ is the number of [predictor variables](#sec-correlationalStudy).
Multiple regression is basically a weighted sum of the [predictor variables](#sec-correlationalStudy), and adding an intercept.
Under the hood, multiple regression seeks to identify the best weight for each [predictor](#sec-correlationalStudy).
The intercept is the expected value of the outcome variable when all of the predictor variables have a value of zero.

@@

When the outcome variable is continuous, [linear regression](#sec-multipleRegressionLinear) is common.
However, there are other types of regression depending on type and distribution of the [outcome variable](#sec-correlationalStudy).
For instance, if the [outcome variable](#sec-correlationalStudy) is binary, [logistic regression](#sec-multipleRegressionLogistic) would be used.
If the [outcome variable](#sec-correlationalStudy) is an ordered categorical variable, [ordinal regression](#sec-multipleRegressionOrdinal) would be used.
If the [outcome variable](#sec-correlationalStudy) is a count, [Poisson](#sec-multipleRegressionPoisson) or [negative binomial](#sec-multipleRegressionNegativeBinomial) regression would be preferable.
If the [outcome variable](#sec-correlationalStudy) is a proportion, [beta regression](#sec-multipleRegressionBeta) would be used.

@@

To evaluate the shape of the association between the [predictor variables](#sec-correlationalStudy) and the [outcome variable](#sec-correlationalStudy), we can examine [scatterplots](#sec-scatterplot) (@fig-linearNonlinearAssociations), residual plots (@fig-residualPlots), marginal model plots (@fig-marginalModelPlots), added-variable plots (@fig-addedVariablePlots), and component-plus-residual plots (@fig-componentPlusResidualPlots).
Residual plots depict the residuals (errors) on the y-axis as a function of the fitted values or a specific [predictor](#sec-correlationalStudy) on the x-axis.
Marginal model plots are basically glorified [scatterplots](#sec-scatterplot) that depict the [outcome variable](#sec-correlationalStudy) (both in terms of observed values and model-fitted values) on the y-axis and the [predictor variables](#sec-correlationalStudy) on the x-axis.
Added-variable plots depict the *unique* association of each [predictor variables](#sec-correlationalStudy) with the [outcome variable](#sec-correlationalStudy) when controlling for all the other [predictor variables](#sec-correlationalStudy) in the model.
Component-plus-residual plots depict partial residuals on the y-axis as a function of each [predictor variable](#sec-correlationalStudy) on the x-axis, where a partial residual for a given [predictor](#sec-correlationalStudy) is the effect of a given predictor (thus controlling for all the other [predictor variables](#sec-correlationalStudy) in the model) plus the residual from the full model.

@@

Let's first examine [descriptive statistics](#sec-descriptiveStatistics) of the predictor and outcome variables.

@@

We can perform a dominance analysis to evaluate the relative importance of the predictors in the regression model.
To perform the dominance analysis, we use the `parameters::dominance_analysis()` function of the `parameters` package [@R-parameters; @Luedecke2020_packages], which leverages the `domir` package [@R-domir].

@@

We can calculate this by substituting in the regression coefficients and predictor values.
Here is the computation:

@@

To address multicollinearity, you can drop a redundant predictor or you can also use [principal component analysis](#sec-pca) or [factor analysis](#sec-factorAnalysis) of the [predictors](#sec-correlationalStudy) to reduce the [predictors](#sec-correlationalStudy) down to a smaller number of meaningful [predictors](#sec-correlationalStudy).
For a meaningful answer regarding [predictors](#sec-correlationalStudy) in a regression framework that is precise and confident, you need a low level of intercorrelation among [predictors](#sec-correlationalStudy), unless you have a very large sample size.
However, if you are merely interested in prediction—and are not interested in interpreting the regression coefficients of individual predictors—multicollinearity poses less of a problem.
For instance, [machine learning](#sec-machineLearning) cares more about achieving the greatest predictive accuracy possible and cares less about explaining which [predictors](#sec-correlationalStudy) are causally related to the [outcome](#sec-correlationalStudy).
So, multicollinearity is less of a concern for [machine learning](#sec-machineLearning) approaches.

@@

- NOT a predictor of the outcome: `0`
- cluster variable: `-2`
- fixed effect of predictor: `1`
- fixed effect and random effect of predictor: `2`
- include cluster mean of predictor in addition to fixed effect of predictor: `3`
- include cluster mean of predictor in addition to fixed effect and random effect of predictor: `4`

>>>>

All statistical analyses follow the same basic structure, as in @eq-analysisBasicStructure [@Brauer2018]:

$$
\text{DATA} = \text{MODEL} + \text{ERROR}
$$ {#eq-analysisBasicStructure}

@@

where $p$ is the number of [predictor variables](#sec-correlationalStudy).
Multiple regression is basically a weighted sum of the [predictor variables](#sec-correlationalStudy), and adding an intercept.
Under the hood, multiple regression seeks to identify the best weight for each [predictor](#sec-correlationalStudy).
The intercept is the expected value of the [outcome variable](#sec-correlationalStudy) when all of the [predictor variables](#sec-correlationalStudy) have a value of zero.

@@

When the [outcome variable](#sec-correlationalStudy) is continuous, [linear regression](#sec-multipleRegressionLinear) is common.
However, there are other types of regression depending on type and distribution of the [outcome variable](#sec-correlationalStudy).
For instance, if the [outcome variable](#sec-correlationalStudy) is binary, [logistic regression](#sec-multipleRegressionLogistic) would be used.
If the [outcome variable](#sec-correlationalStudy) is an ordered categorical variable, [ordinal regression](#sec-multipleRegressionOrdinal) would be used.
If the [outcome variable](#sec-correlationalStudy) is a count, [Poisson](#sec-multipleRegressionPoisson) or [negative binomial](#sec-multipleRegressionNegativeBinomial) regression would be preferable.
If the [outcome variable](#sec-correlationalStudy) is a proportion, [beta regression](#sec-multipleRegressionBeta) would be used.

@@

To evaluate the shape of the association between the [predictor variables](#sec-correlationalStudy) and the [outcome variable](#sec-correlationalStudy), we can examine [scatterplots](#sec-scatterplot) (@fig-linearNonlinearAssociations), residual plots (@fig-residualPlots), marginal model plots (@fig-marginalModelPlots), added-variable plots (@fig-addedVariablePlots), and component-plus-residual plots (@fig-componentPlusResidualPlots).
Residual plots depict the residuals (errors) on the y-axis as a function of the fitted values or a specific [predictor](#sec-correlationalStudy) on the x-axis.
Marginal model plots are basically glorified [scatterplots](#sec-scatterplot) that depict the [outcome variable](#sec-correlationalStudy) (both in terms of observed values and model-fitted values) on the y-axis and the [predictor variables](#sec-correlationalStudy) on the x-axis.
Added-variable plots depict the *unique* association of each [predictor variables](#sec-correlationalStudy) with the [outcome variable](#sec-correlationalStudy) when controlling for all the other [predictor variables](#sec-correlationalStudy) in the model.
Component-plus-residual plots depict partial residuals on the y-axis as a function of each [predictor variable](#sec-correlationalStudy) on the x-axis, where a partial residual for a given [predictor](#sec-correlationalStudy) is the effect of a given [predictor](#sec-correlationalStudy) (thus controlling for all the other [predictor variables](#sec-correlationalStudy) in the model) plus the residual from the full model.

@@

Let's first examine [descriptive statistics](#sec-descriptiveStatistics) of the [predictor](#sec-correlationalStudy) and [outcome](#sec-correlationalStudy) variables.

@@

We can perform a dominance analysis to evaluate the relative importance of the [predictors](#sec-correlationalStudy) in the regression model.
To perform the dominance analysis, we use the `parameters::dominance_analysis()` function of the `parameters` package [@R-parameters; @Luedecke2020_packages], which leverages the `domir` package [@R-domir].

@@

We can calculate this by substituting in the regression coefficients and [predictor](#sec-correlationalStudy) values.
Here is the computation:

@@

To address multicollinearity, you can drop a redundant [predictor](#sec-correlationalStudy) or you can also use [principal component analysis](#sec-pca) or [factor analysis](#sec-factorAnalysis) of the [predictors](#sec-correlationalStudy) to reduce the [predictors](#sec-correlationalStudy) down to a smaller number of meaningful [predictors](#sec-correlationalStudy).
For a meaningful answer regarding [predictors](#sec-correlationalStudy) in a regression framework that is precise and confident, you need a low level of intercorrelation among [predictors](#sec-correlationalStudy), unless you have a very large sample size.
However, if you are merely interested in prediction—and are not interested in interpreting the regression coefficients of individual predictors—multicollinearity poses less of a problem.
For instance, [machine learning](#sec-machineLearning) cares more about achieving the greatest predictive accuracy possible and cares less about explaining which [predictors](#sec-correlationalStudy) are causally related to the [outcome](#sec-correlationalStudy).
So, multicollinearity is less of a concern for [machine learning](#sec-machineLearning) approaches.

@@

- NOT a [predictor](#sec-correlationalStudy) of the outcome: `0`
- [cluster variable](#sec-mixedModelDataStructure): `-2`
- [fixed effect](#sec-mixedModelsOverview) of [predictor](#sec-correlationalStudy): `1`
- [fixed effect](#sec-mixedModelsOverview) and [random effect](#sec-mixedModelsOverview) of [predictor](#sec-correlationalStudy): `2`
- include cluster mean of [predictor](#sec-correlationalStudy) in addition to [fixed effect](#sec-mixedModelsOverview) of [predictor](#sec-correlationalStudy): `3`
- include cluster mean of [predictor](#sec-correlationalStudy) in addition to [fixed effect](#sec-mixedModelsOverview) and [random effect](#sec-mixedModelsOverview) of [predictor](#sec-correlationalStudy): `4`
